---
title: "Penguins ANOVA"
author: "Eleonore Durand"
date: "November 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Load packages and read-in penguins.csv data

```{r, include = FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")


```


Our question: 
Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, and Adelie)?

ANOVA here makes sense because comparing multiple groups. Note, ANOVA can only be used if we satisfy some assumptions. 

Step 2. Do some exploring

New functions: with + table for easy counting over multiple factors

Update species name using case_when

```{r}

counts <- with(penguins, table(species, sex)) #creates a table of counts of each species of penguin by species and sex

penguins2 <- penguins %>% 
  mutate(common_name = 
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap", 
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
           )) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")



```

- Exploratory graphs
- Summary statistics 
- Levene's tests for equal variances
- ANOVA

