---
title: "Penguins ANOVA"
author: "Eleonore Durand"
date: "November 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Load packages and read-in penguins.csv data

```{r, include = FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")


```


Our question: 
Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, and Adelie)?

ANOVA here makes sense because comparing multiple groups. Note, ANOVA can only be used if we satisfy some assumptions. 

Step 2. Do some exploring

New functions: with + table for easy counting over multiple factors

Update species name using case_when

```{r}

counts <- with(penguins, table(species, sex)) #creates a table of counts of each species of penguin by species and sex

penguins2 <- penguins %>% 
  mutate(common_name = 
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap", 
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
           )) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")



```

- Exploratory graphs
- Summary statistics 
- Levene's tests for equal variances
- one-way ANOVA
- post-hoc testing 

```{r}
hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) + 
  facet_wrap(~common_name) +
  theme_classic()

hists 

qqplots <- ggplot(penguins2, aes(sample = flipper_length)) + 
  geom_qq(aes(color = common_name)) +
  facet_wrap(~common_name) +
  theme_classic()

qqplots


# Levene's test for equal variances (>2 groups.... if you have 2 groups, use the F-test for equal variances)

#We'll use the function leveneTest from the 'car' package

# H0: Variances are equal
# HA: Variances are unequal

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)
penguin_levene

# We retain the null hypothesis of equal variances (p = 0.06)

#Use group_by and summarize to do the variances for each group 

var_penguin <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )
var_penguin
```


#3. One-way ANOVA:

single factor: common_name (species)
Number of levels in the factor : 3 (Chinstrap, Gentoo, Adelie)
Random variable: flipper_length (mm)

Is there a significatn difference in mean flipper length (mm) for female chinstrap, Gentoo, and Adelie penguins?


```{r}
#H0: Mean flipper lengths across all species are equal
#HA: At LEAST two means differ significantly 

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov) #summary of penguin_aov gives you more information because it didnt otherwise

#If we take samples from all 3 groups that have equal means, then there is a 0% chance that we could have found means that were at least this different or more, by random chance. 

#At least two samples were taken from populations with different means. There is a significant difference for flipper length across the three species (Adelie, Chinstrap and Gentoo) (F(df) = statistic, p<0.001, $\alpha$= 0.005 .

#Which ones are different? All three different from each other? Or something else?

#Post-hoc testing using Tukey's HSD

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

```


Mean flipper lengths for Adelie, Gentoo, and Chinstrap penguins are all significantly different (F(df) = statistic, p<0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha = 0.05$)